Curvature approximation on surfaces -- A Discrete Exterior Calculus Approach

### 2014-11-28 article/main.tex
...

@@

    \subsubsection{The discrete gradient}
      The gradient of a scalar valued function \( f \) can be obtained by rising the indices of its exterior derivative, 
      in other words \( \nabla f = \left( \exd f \right)^{\sharp} \).
      Unfortunately, the sharp operator \( \sharp \) depends on the (unknown) metric of the manifold. 
      In \cite{hirani} it is suggested to use this discrete sharp operator for such a exact \mh{1}{form} \( \exd f \) to get a discrete
      gradient:
      \begin{align}
        \label{eqGradientHeine}
        \left\langle \left( \exd f \right)^{\sharp} , \star\sigma^{2}\right\rangle
          &:= \sum_{\sigma^{0}\prec\sigma^{2}} \left( f(\sigma^{0}) - f(w) \right) \nabla\Phi_{\sigma^{0}}^{\sigma^{2}} \formkomma
      \end{align}
      with a arbitrary vertex \( w\prec\sigma^{2} \) and 
      \( \Phi_{\sigma^{0}}^{\sigma^{2}} \) is the linear test function (hat function) for the vertex \( \sigma^{0} \) 
      restricted to \( \sigma^{2} \).

>>>>

  \begin{figure}
    \centering
    \includegraphics[width=.45\textwidth]{bilder/bunnyCurvature.jpg}
    \caption{blub}
    \label{figBunnyCurvature}
  \end{figure}

@@

    \subsubsection{The discrete gradient}
      The gradient of a scalar valued function \( f \) can be obtained by rising the indices of its exterior derivative, 
      in other words \( \nabla f = \left( \exd f \right)^{\sharp} \).
      Unfortunately, the sharp operator \( \sharp \) depends on the (unknown) metric of the manifold. 
      In \cite{hirani} it is suggested to use the following discrete sharp operator for such a exact \mh{1}{form} \( \exd f \) to get a discrete
      gradient:
      \begin{align}
        \label{eqGradientHeine}
        \left\langle \left( \exd f \right)^{\sharp} , \star\sigma^{2}\right\rangle
          &:= \sum_{\sigma^{0}\prec\sigma^{2}} \left( f(\sigma^{0}) - f(w) \right) \nabla\Phi_{\sigma^{0}}^{\sigma^{2}} \formkomma
      \end{align}
      with a arbitrary vertex \( w\prec\sigma^{2} \) and 
      \( \Phi_{\sigma^{0}}^{\sigma^{2}} \) is the linear test function (hat function) for the vertex \( \sigma^{0} \) 
      restricted to \( \sigma^{2} \). 
      \eqref{eqGradientHeine} gives us a tangential vector of the polytop \( |K| \) on the circumcenter of the triangle \( \sigma^{2} \).
      But in our setting we want to use a gradient on a vertex of the primal mesh, resp. interpreted on the associated Voronoi cell.
      A simple way to get such a discrete gradient, is to average \eqref{eqGradientHeine} over all dual vertices of the Voronoi cell
      related to the primal vertex \( v \).
      \begin{align} \label{eqGradientAverage}
        \left\langle *\overline\nabla f, \star v \right\rangle
            = \sum_{\sigma^{2}\succ v} \left| \star v \cap \sigma^{2} \right|
                 \sum_{\sigma^{0}\prec\sigma^{2}} \left( f(\sigma^{0}) - f(v) \right) \nabla\Phi_{\sigma^{0}}^{\sigma^{2}}
      \end{align}
      The drawback of the discrete average gradient, which we will use in section \ref{secWeingarten}, is that \( |K| \) has no tangential
      space on its vertices, because the polytop is not continuous differentiable in \( v \).
      This makes the situation for a numerical analysis slightly difficult.
      But comparing calculations for several manifolds and functions indicate the consistence of \eqref{eqGradientAverage} to the smooth
      gradient, 
      if the triangulation is fine enough.
